How to select predictive models for decision-making or causal inference?

### 2023-12-16 how_to_select_causal_models.tex
iter

\onecolumn
\maketitle
\begin{abstract}
    \textbf{Objective:}
    As predictive models --\eg from machine learning-- give likely
    outcomes, they may be used to reason on the effect of an intervention, a
    causal-inference task. The increasing complexity of health data has opened
    the door to a plethora of models, but also the Pandora box of model
    selection: which of these models yield the most valid causal estimates? 
    We study such model selection procedures in practical settings:
    finite samples settings and without well-specification assumption.
    \textbf{Methods:}
    Elaborate risks build proxies of the causal error using ``nuisance''
    re-weighting to compute it on the observed data. We evaluate whether
    empirically estimated nuisances, which are necessarily noisy,
    add noise to model selection. Drawing from an extensive empirical
    study based on a simulation and three healthcare datasets
    based on real covariates, we compare different metrics for causal model
    selection.\\
    \textbf{Results:} Among all metrics the classical mean squared error (ie.
    factual $\mu\mathrm{-risk}$) is worse and re-weighting it with propensity score
    ($\mu\mathrm{-risk}_{IPW}$) does not bring much improvements. The
    $R\text{-risk}$, which includes a model of mean outcome and propensity
    scores, leads to the best performances.\\
    \textbf{Conclusions:} We outline that evaluating models for causal
    reasoning requires different procedures than standard predictive
    settings: using the so-called $R\text{-risk}$ with nuisance
    corrections computed with flexible estimators.
\end{abstract}
\keywords{Model Selection, Treatment Effect, G-formula, Observational Study, Machine Learning, }
\clearpage

@@

Progress in machine learning brings predictive models to new health data
\cite{beam2018big,rajkomar2019machine}, with the promise of precision medicine.
Predictive models are increaslingly used for image analysis
\cite{khojaste2022deep,zhang2019radiological,yala2019deep,shen2019deep,nassif2022breast},
routinely-collected patient records
\cite{mooney2018bigdata,desai2020comparison,simon2018predicting} or clinical
notes \cite{horng2017creating,wang2020prediction,spasic2020clinical}. Data may
be difficult to control and model, but the accuracy of the prediction can be
verified on left-out data
\cite{altman2009prognosis,poldrack2020establishment,varoquaux2022evaluating}.
Given a model predicting a health outcome, precision medicine would like it to
guide decisions: will an individual benefit from an intervention such as surgery
\cite{fontana2019can}? Contrasting predictions with and without the treatment
gives an answer, but statistical validity requires causal inference
\cite{snowden_implementation_2011,sperrin2019explicit,blakely2020reflection}.

>>>>

\onecolumn
\maketitle
\begin{abstract}\noindent
    \textbf{Objective:}
    To evaluate which procedure selects the predictive model most
    trustworthy to reason on the effect of an intervention and support
    decision making.\\
    %As predictive models --\eg from machine learning-- give likely
    %outcomes, they may be used to, a
    %causal-inference task. The increasing complexity of health data has opened
    %the door to a plethora of models, but also the Pandora box of model
    %selection: which of these models yield the most valid causal estimates? 
    \textbf{Methods:}
    We study such a variety of model selection procedures in practical settings:
    finite samples settings and without well-specification assumption.
    Elaborate risks build proxies of the causal error using ``nuisance''
    re-weighting to compute it on the observed data. We evaluate whether
    empirically estimated nuisances, which are necessarily noisy,
    add noise to model selection. Drawing from an extensive empirical
    study based on a simulation and three healthcare datasets
    based on real covariates, we compare different metrics for causal model
    selection.\\
    \textbf{Results:} Among all metrics the classical mean squared error (ie.
    factual $\mu\mathrm{-risk}$) is worse and re-weighting it with propensity score
    ($\mu\mathrm{-risk}_{IPW}$) does not bring much improvements. The
    $R\text{-risk}$, which includes a model of mean outcome and propensity
    scores, leads to the best performances.\\
    \textbf{Conclusions:} We outline that evaluating models for causal
    reasoning requires different procedures than standard predictive
    settings: using the so-called $R\text{-risk}$ with nuisance
    corrections computed with flexible estimators.
\end{abstract}
\keywords{Model Selection, Treatment Effect, G-formula, Observational Study, Machine Learning, }
\clearpage

@@

Progress in machine learning brings predictive models to new health data
\cite{beam2018big,rajkomar2019machine}, with the promise of precision medicine.
Predictive models are increasingly used for image analysis
\cite{khojaste2022deep,zhang2019radiological,yala2019deep,shen2019deep,nassif2022breast},
patient records
\cite{mooney2018bigdata,desai2020comparison,simon2018predicting} or clinical
notes \cite{horng2017creating,wang2020prediction,spasic2020clinical}. Data may
be difficult to control and model, but the accuracy of the prediction can be
verified on left-out data
\cite{altman2009prognosis,poldrack2020establishment,varoquaux2022evaluating}.
Given a model predicting a health outcome, precision medicine would like it to
guide decisions: will an individual benefit from an intervention such as surgery
\cite{fontana2019can}? Contrasting predictions with and without the treatment
gives an answer, but statistical validity requires causal inference
\cite{snowden_implementation_2011,sperrin2019explicit,blakely2020reflection}.
